### 2021-03-12 02_abstract/abstract.tex
Restructured thesis and removed unnecessary parts
Restructured thesis and added more PDF bookmarks
Optimized chapter 1, 2 and 7

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Abstract (English)                              %
%=============================================================================%

\chapter*{Abstract (English version)}

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
numerical simulations of the complex physical, biological and chemical
processes of the Earth system. Of particular relevance are projections of the
future climate, which enable an assessment of future climate change. Due to the
chaotic nature of the climate system, unknowns in future emission pathways and
uncertainties in the climate models, projections of the future climate are
associated with large uncertainties. In this thesis, we analyze future climate
projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and reduce
corresponding uncertainties with observations.

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, the equilibrium
temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, we find a \acfAbstract{MMM} and multi-model
range of $3.74 \unit{K}$ and $\rangeunit{1.8}{5.6}{K}$, respectively. These
values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
third of the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely
\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR}
(\acs{AR}5) of the \acl{IPCC} from 2013. Similarly, the transient response of
the climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean, which reduces
the strong negative shortwave cloud phase change feedback in this region that
is present in climate models from previous \acs{CMIP} generations.

To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, we
systematically analyze eleven published emergent constraints on \acs{ECS} that
have mostly been derived from models participating in \acs{CMIP}5, the
predecessor generation of \acs{CMIP}6. Emergent constraints are potentially
promising approaches to reduce uncertainties in climate model projections by
combining observations and output from \acsp{ESM}. The focus of this study is
on testing if these emergent constraints hold for \acsp{ESM} participating
\acs{CMIP}6. Since none of the emergent constraints considered here have been
derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
cross-checking of the emergent constraints on a new model ensemble. The
application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
skill and statistical significance of the emergent relationship for nearly all
constraints, with this decrease being large in many cases. Consequently, the
size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
This is likely related to the increased multi-model spread of \acs{ECS} in
\acs{CMIP}6, but may in some cases also be due to spurious statistical
relationships or a too small number of models in the ensemble that the emergent
constraint was originally derived from. The corresponding best estimates given
by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
$12 \unit{\%}$ on average. This can be at least partly explained by the
increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
corresponding change in the constraint predictors, suggesting the emergence of
new feedback processes rather than changes in strength of those previously
dominant. Our results support previous studies concluding that emergent
constraints should be based on an independently verifiable physical mechanism
and that process-based emergent constraints on \acs{ECS} should rather be
thought of as constraints for the process or feedback they are actually
targeting.

To overcome these issues of single-process-oriented emergent constraints, we
introduce an alternative approach based on \acfAbstract{ML}. Since this new
technique relies on a large number of data points in order to train the
\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
\acs{TCR} is not an appropriate target variable. As an alternative, we study
projections of the future \acfAbstract{GPP}, which is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \acs{CO2}
from the atmosphere. In this analysis, we combine an existing emergent
constraint on \acs{CO2} fertilization with a \acs{ML} approach to constrain the
spatial variations of multi-model \acs{GPP} projections. In the first step of
our two-step approach, we use observed changes in the \acs{CO2} seasonal cycle
at \acl{KUM} to constrain the global mean \acs{GPP} at the end of the \nth{21}
century (\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM}
participating in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
the unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second
step, we use a \acs{ML} model to constrain gridded future absolute \acs{GPP}
and gridded fractional \acs{GPP} change in two independent approaches. For
this, observational data is fed into the \acs{ML} algorithm that has been
trained on \acs{CMIP}5 data to learn relationships between present-day
physically-relevant diagnostics and the target variable. In a
leave-one-model-out \acl{CV} approach, the \acs{ML} model shows superior
performance to the \acs{CMIP}5 \acs{MMM}. Our approach predicts an increased
\acs{GPP} change in northern high latitudes compared to regions closer to the
equator.

>>>>

%=============================================================================%
%                            Doctoral Dissertation                            %
%                               Manuel Schlund                                %
%                                  (c) 2021                                   %
%=============================================================================%
%                                  Abstract                                   %
%=============================================================================%

\chapter{\abstractname}

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
numerical simulations of the complex physical, biological and chemical
processes of the Earth system. Of particular relevance are projections of the
future climate, which enable an assessment of future climate change. Due to the
chaotic nature of the climate system, unknowns in future emission pathways and
uncertainties in the climate models, projections of the future climate are
associated with large uncertainties. The main focus of this thesis is the
analysis of future climate projections from \acsp{ESM} participating in the
\acfAbstract{CMIP} and the reduction of corresponding uncertainties with
observations.

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, which is defined as
the equilibrium temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR} (\acs{AR}5)
of the \acl{IPCC} from 2013. Similarly, the transient response of the climate
system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR}, shows an
inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is again
higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean, which reduces
the strong negative shortwave cloud phase change feedback in this region that
is present in climate models from previous \acs{CMIP} generations.

To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models,
eleven published emergent constraints on \acs{ECS} (mostly derived from models
participating in \acs{CMIP}5, the predecessor generation of \acs{CMIP}6) are
systematically analyzed. Emergent constraints are potentially promising
approaches to reduce uncertainties in climate model projections by combining
observations and output from \acsp{ESM}. The focus of this analysis is on
testing if these emergent constraints hold for \acsp{ESM} participating
\acs{CMIP}6. Since none of the emergent constraints considered here have been
derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
cross-checking of the emergent constraints on a new model ensemble. The
application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
skill and statistical significance of the emergent relationship for nearly all
constraints, with this decrease being large in many cases. Consequently, the
size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
This is likely related to the increased multi-model spread of \acs{ECS} in
\acs{CMIP}6, but may in some cases also be due to spurious statistical
relationships or a too small number of models in the ensemble that the emergent
constraint was originally derived from. The corresponding best estimates given
by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
$12 \unit{\%}$ on average. This can be at least partly explained by the
increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
corresponding change in the constraint predictors, suggesting the emergence of
new feedback processes rather than changes in strength of those previously
dominant. The results support previous studies concluding that emergent
constraints should be based on an independently verifiable physical mechanism
and that process-based emergent constraints on \acs{ECS} should rather be
thought of as constraints for the process or feedback they are actually
targeting.

To overcome these issues of single-process-oriented emergent constraints, an
alternative approach based on \acfAbstract{ML} is introduced. Since this new
technique relies on a large number of data points in order to train the
\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
\acs{TCR} is not an appropriate target variable. Therefore, \acfAbstract{GPP}
as a process that contributes to climate sensitivity is studied as an
alternative. \acs{GPP} is the largest flux of the terrestrial carbon uptake and
slows down global warming by removing \acs{CO2} from the atmosphere. In this
analysis, an existing emergent constraint on \acs{CO2} fertilization is
combined with a \acs{ML} approach to constrain the spatial variations of
multi-model \acs{GPP} projections. In the first step of the two-step approach,
observed changes in the \acs{CO2} seasonal cycle at \acl{KUM} are used to
constrain the global mean \acs{GPP} at the end of the \nth{21} century
(\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM} participating
in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the
unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second step, a
\acs{ML} model is used to constrain gridded future absolute \acs{GPP} and
gridded fractional \acs{GPP} change in two independent approaches. For this,
observational data is fed into the \acs{ML} algorithm that has been trained on
\acs{CMIP}5 data to learn relationships between present-day physically-relevant
diagnostics and the target variable. In a leave-one-model-out \acl{CV}
approach, the \acs{ML} model shows superior performance to the \acs{CMIP}5
\acs{MMM}. The new approach predicts a higher \acs{GPP} increase in high
latitudes and a lower \acs{GPP} increase in regions closer to the equator.

### 2021-03-12 ch01_introduction/introduction.tex
Restructured thesis and added more PDF bookmarks
Optimized chapter 1, 2 and 7

Climate change is one of the greatest challenges for humankind today with
far-reaching consequences for all human beings on planet Earth. The warming of
the climate system is \enquote{unequivocal} and \enquote{many of the observed
  changes are unprecedented over decades to millennia} \autocite{IPCC2014}. The
changing climate increases the \enquote{likelihood of severe, pervasive and
  irreversible impacts for people, species and ecosystems} with \enquote{mostly
  negative impacts for biodiversity, ecosystem services and economic
  development} and amplifies \enquote{risks for livelihoods and for food and
  human security} \autocite{IPCC2014}. Potential drivers for climate change
are all natural and anthropogenic substances and processes that may alter the
Earth's energy budget. The human influence on the climate system is clear: Ever
increasing emissions of \acp{GHG} since the end of the pre-industrial era
largely driven by economic and population growth led to atmospheric
concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
\enquote{unprecedented in at least the last 800000 years} \autocite{IPCC2014}.
The effects of these \ac{GHG} emissions and other anthropogenic drivers have
been \enquote{detected throughout the climate system and are extremely likely
  to have been the dominant cause of the observed warming since the
  mid-\nth{20} century} \autocite{IPCC2014}. The greenhouse effect is based on
the optical properties of the \acp{GHG}: While the atmosphere is mostly
transparent for the incoming solar (shortwave) radiation, the outgoing
infrared (longwave) radiation that is reflected from the planet's surface is
partly absorbed by the corresponding \ac{GHG} molecules through an excitation
of their corresponding vibrational modes and re-emitted isotropically. This
process traps energy near the surface and leads to a warming of the Earth's
surface and the lower atmosphere. The individual impact of the different
drivers of climate change measured with the so-called \emph{radiative
forcing}, which quantifies the change in energy fluxes caused by changes in
these drivers relative to pre-industrial conditions \autocite{IPCC2013}. The
total radiative forcing is positive (which corresponds to a warming) and its
largest contribution is caused by the increase in the atmospheric
concentration of \ac{CO2} due to fossil fuel emissions since the year 1750
\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
of the terrestrial biosphere due to photosynthesis. Since this land carbon
uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions in
today's climate \autocite{Friedlingstein2020}, this process substantially slows
down global warming. Other important anthropogenic drivers of climate change
are the emission of aerosols and land use/land cover changes. Apart from
anthropogenic drivers, there are also natural processes which impact the
climate system like changes in the solar activity or volcanic eruptions.
However, there is clear evidence that these natural drives alone cannot explain
the observed climate change \autocite{Haustein2017}.

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes provide a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} will be used in the upcoming Sixth \acl{AR} (\acs{AR}6) of
the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been successfully evaluated in the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\acp{ESM} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. A state-of-the-art technique for this is the \emph{emergent
  constraints} method, which uses a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is the weighting of climate
models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

@@

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). A complete
list of these is given on \cpageref{ch:bib:own}. Wherever material from these
studies is presented in this thesis, the pronoun \enquote{we} is used to
increase readability by avoiding the passive voice and to acknowledge all
involved contributors. However, unless stated otherwise, all contents from
these publications (text, figures and tables) shown in this thesis originate
from the author of this thesis. A detailed list of contributions to these
studies (including their full reference) is given in the corresponding
\namecrefs{ch:03:esmvaltool}.

>>>>

Climate change is one of the greatest challenges for humankind today with
far-reaching consequences for all human beings on planet Earth. The warming of
the climate system is \enquote{unequivocal} and \enquote{many of the observed
  changes are unprecedented over decades to millennia} \autocite{IPCC2014}. The
changing climate increases the \enquote{likelihood of severe, pervasive and
  irreversible impacts for people, species and ecosystems} with \enquote{mostly
  negative impacts for biodiversity, ecosystem services and economic
  development} and amplifies \enquote{risks for livelihoods and for food and
human security} \autocite{IPCC2014}. Potential drivers for climate change are
all natural and anthropogenic substances and processes that may alter the
Earth's energy budget. The human influence on the climate system is clear:
Ever increasing emissions of \acp{GHG} since the end of the pre-industrial
era largely driven by economic and population growth led to atmospheric
concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
\enquote{unprecedented in at least the last 800000 years}
\autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
anthropogenic drivers have been \enquote{detected throughout the climate
system and are extremely likely to have been the dominant cause of the
observed warming since the mid-\nth{20} century} \autocite{IPCC2014}. The
greenhouse effect is based on the optical properties of the \acp{GHG}: While
the atmosphere is mostly transparent for the incoming solar (shortwave)
radiation, the outgoing infrared (longwave) radiation that is reflected from
the planet's surface is partly absorbed by the corresponding \ac{GHG}
molecules through an excitation of their corresponding vibrational modes and
re-emitted isotropically. This process traps energy near the surface and
leads to a warming of the Earth's surface and the lower atmosphere. The
individual impact of the different drivers of climate change measured with
the so-called \emph{radiative forcing}, which quantifies the change in energy
fluxes caused by changes in these drivers relative to pre-industrial
conditions \autocite{IPCC2013}. The total radiative forcing is positive
(which corresponds to a warming) and its largest contribution is caused by
the increase in the atmospheric concentration of \ac{CO2} due to fossil fuel
emissions since the year 1750 \autocite{IPCC2014}. Apart from their physical
warming effect on the climate, carbon-based \acp{GHG} like \ac{CO2} also
directly influence the global carbon cycle, an important biogeochemical cycle
of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}}, which
describes the carbon uptake of the terrestrial biosphere due to photosynthesis.
Since this land carbon uptake absorbs about $30 \unit{\%}$ of the anthropogenic
\ac{CO2} emissions in today's climate \autocite{Friedlingstein2020}, this
process substantially slows down global warming and directly contributes to the
magnitude of the climate sensitivity. Other important anthropogenic drivers of
climate change are the emission of aerosols and land use/land cover changes.
Apart from anthropogenic drivers, there are also natural processes which impact
the climate system like changes in the solar activity or volcanic eruptions.
However, there is clear evidence that these natural drivers alone cannot
explain the observed climate change \autocite{Haustein2017}.

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes provide a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} will be used in the upcoming Sixth \acl{AR} (\acs{AR}6) of
the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been successfully analyzed in the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\acp{ESM} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with large ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. A state-of-the-art technique for this is the \emph{emergent
  constraints} method, which uses a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is the weighting of climate
models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

@@

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). A complete
list of these is given on \cpageref{ch:bib:own}. Wherever material from these
studies is presented in this thesis, the pronoun \enquote{we} is used to
increase readability by avoiding the passive voice and to acknowledge all
involved contributors. However, unless stated otherwise, all contents from
these publications (text, figures and tables) shown in this thesis originate
from the author of this thesis. A detailed list of contributions to these
studies is given in the corresponding \namecrefs{ch:03:esmvaltool}.
